---
title: "Topic 16: Inference Practice (Part A)"
tutorial:
  id: "InferencePractice.Topic16a.AppliedStats"
  version: 1.0
output: 
  learnr::tutorial:
    progressive: TRUE
    allow_skip: TRUE
    
runtime: shiny_prerendered
description: "No new content is introduced in this notebook. The original Topic 16notebook simply contains twelve practice problems involving inference on mean(s), inference on proportion(s), or estimating sample sizes. This notebook only contains the first four problems so that the RAM overflow on rebuild problem is bypassed."
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(learnr)
library(gradethis)


tutorial_options(exercise.timelimit = 60, exercise.checker = gradethis::grade_learnr)

```

## Inference Practice

###

In this workbook, you'll have an opportunity to practice a set of problems. I'm still providing you with a set of guided questions for each problem. Think about the guiding questions and steps necessary to complete each task so that you will be prepared to approach similar problems on your own, without this scaffolding. On the associated homework problems you'll need to come up with these steps on your own. This workbook should prepare you for that.

###

For convenience, here is the [*standard errror decision tree*](https://agmath.github.io/SiteFiles/StdErrorDecisionTree.pdf){target="_blank"}, the general [*strategy for confidence intervals*](https://agmath.github.io/SiteFiles/CIgeneral.pdf){target="_blank"}, and the general [*strategy for hypothesis tests*](https://agmath.github.io/SiteFiles/HTgeneral.pdf){target="_blank"}. For any problems involving hypothesis tests, assume the $\alpha = 0.05$ level of significance unless otherwise stated.

### 

> **Note:** There are no hints included in this notebook. You should use the resources linked above and your previous notebooks to help you. These first four problems include links to fully worked solutions, but you'll be on your own for the remaining problems.

## Problem 1

**Problem 1:** The distribution of the number of eggs laid by a certain species of hen during their breeding period is 35 eggs with a standard deviation of 18.2. Suppose a group of researchers randomly samples 45 hens of this species, counts the number of eggs laid during their breeding period, and records a sample mean of 37 eggs. Find the probability of observing a sample of 45 hens whose mean number of eggs laid during the breeding period is at least 37. (In case of emergency, [see worked solution here](www/SampleInferenceProbSol.pdf){target="_blank"})

```{r prob1-1, echo = FALSE}
question_radio(
  "To answer the question as asked, we should",
  answer("Construct a confidence interval"),
  answer("Conduct a hypothesis test"),
  answer("Compute a probability", correct = TRUE),
  answer("Find a required sample size"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob1-2, echo = FALSE}
question_radio(
  "How many groups are samples in this application representing?",
  answer("The sample was taken from a single population.", correct = TRUE),
  answer("Sample observations were collected from two populations."),
  answer("Sample observatons were collected from more than two populations."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

```{r prob1-3, echo = FALSE}
question_radio(
  "Do we know the population standard deviation for number of eggs laid?",
  answer("Yes, the population standard deviation is known.", correct = TRUE),
  answer("No, the population standard deviation is not known."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob1-4, echo = FALSE}
quiz(
  question_radio(
    "How should the standard error of the sampling distribution be computed?",
    answer("$\\displaystyle{S_E = \\sigma/\\sqrt{n}}$", correct = TRUE),
    answer("$\\displaystyle{S_E = s/\\sqrt{n}}$"),
    answer("$\\displaystyle{S_E = \\sigma_{\\text{diff}}/n_{\\text{diff}}}$"),
    answer("$\\displaystyle{S_E = s_{\\text{diff}}/\\sqrt{n_{\\text{diff}}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{\\sigma_1^2}{n_1} + \\frac{\\sigma_2^2}{n_2}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{s_1^2}{n_1} + \\frac{s_2^2}{n_2}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{p\\left(1 - p\\right)}{n}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{p_1\\left(1 - p_1\\right)}{n_1} + \\frac{p_2\\left(1 - p_2\\right)}{n_2}}}$"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "Which distribution best describes the sampling distribution?",
    answer("A normal distribution", correct = TRUE),
    answer("A $t$-distribution with $n - 1$ degrees of freedom."),
    answer("A $t$-distribution with $\\min\\{n_1 - 1, n_2 - 1\\}$ degrees of freedom."),
    answer("A $t$-distribution with $n_{\\text{diff}} - 1$ degrees of freedom."),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

Use the code block below to compute the probability of observing a random sample of 45 hens from this species which lays an average of at least 37 eggs.

```{r prob1-5, exercise = TRUE}

```

```{r prob1-5-check}
grade_result(
  pass_if(~ (abs(.result - (1 - pnorm(37, 35, 18.2/sqrt(45)))) < 0.0005))
)
```

## Problem 2

**Problem 2:** A recent study looked at average ages of male and female inmates on Death Row. The random sample of 9 women and 14 men on death row resulted in the following: 

<center>
$\begin{array}{|c|c|c|} \hline & \text{Men} & \text{Women} \\ \hline
\bar{x} & 39 & 44\\ \hline
s & 4.5 & 6.2 \\ \hline \end{array}$
</center>

Conduct a hypothesis test to determine whether there is evidence to suggest that the men on Death Row are, on average, younger than women on Death Row. (In case of emergency, [see worked solution here](www/SampleInferenceDeathRow.pdf){target="_blank"})

###

```{r prob2-1, echo = FALSE}
question_radio(
  "To answer the question as asked, we should",
  answer("Construct a confidence interval"),
  answer("Conduct a hypothesis test", correct = TRUE),
  answer("Compute a probability"),
  answer("Find a required sample size"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob2-2, echo = FALSE}
question_radio(
  "What is the level of significance associated with this test?",
  answer("$\\alpha = 0.01$"),
  answer("$\\alpha = 0.10$"),
  answer("$\\alpha = 0.05$", correct = TRUE),
  answer("The $p$-value"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob2-3, echo = FALSE}
question_radio(
  "Does this hypothesis test involve testing a statement about the mean ( $\\mu$ ), a proportion ( $p$ ), or something else?",
  answer("One or more means", correct = TRUE),
  answer("One or more proportions"),
  answer("Something else altogether"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob2-4, echo = FALSE}
question_radio(
  "How many groups are being compared in this test?",
  answer("The test involves only a single group"),
  answer("The test compares two groups", correct = TRUE),
  answer("The test compares more than two groups"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob2-5, echo = FALSE}
question_radio(
  "Which of the following are the hypotheses associated with this test?",
  answer("$\\begin{array}{ll} H_0: & \\mu_{\\text{men}} - \\mu_{\\text{women}} = 0\\\\ H_a: & \\mu_{\\text{men}} - \\mu_{\\text{women}} \\neq 0\\end{array}$"),
  answer("$\\begin{array}{ll} H_0: & \\mu_{\\text{men}} - \\mu_{\\text{women}} = 0\\\\ H_a: & \\mu_{\\text{men}} - \\mu_{\\text{women}} > 0\\end{array}$"),
  answer("$\\begin{array}{ll} H_0: & \\mu_{\\text{men}} - \\mu_{\\text{women}} = 0\\\\ H_a: & \\mu_{\\text{men}} - \\mu_{\\text{women}} < 0\\end{array}$", correct = TRUE),
  answer("$\\begin{array}{ll} H_0: & \\mu_{\\text{men}} \\neq \\mu_{\\text{women}}\\\\ H_a: & \\mu_{\\text{men}} = \\mu_{\\text{women}}\\end{array}$"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob2-6, echo = FALSE}
question_radio(
  "Do we know the population standard deviation(s), $\\sigma$, for age in each group?",
  answer("Yes, the population standard deviations are known.", message = "Oops. We are presented with a standard deviation in ages for each group, but those standard deviations belong only to the ages measured in each sample, not the population."),
  answer("No, the population standard deviations are not known.", correct = TRUE),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob2-7, echo = FALSE}
question_radio(
  "Are the observations in the two groups (males on Death Row and females on Death Row) paired?",
  answer("No. There is no reason to suggest that the prisoners are paired.", correct = TRUE),
  answer("Yes. Every male prisoner can be naturally paired to a female prisoner."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob2-8, echo = FALSE}
quiz(
  question_radio(
    "Which standard error formula should be used in the computation of the test statistic?",
    answer("$\\displaystyle{S_E = \\sigma/\\sqrt{n}}$"),
    answer("$\\displaystyle{S_E = s/\\sqrt{n}}$"),
    answer("$\\displaystyle{S_E = \\sigma_{\\text{diff}}/n_{\\text{diff}}}$"),
    answer("$\\displaystyle{S_E = s_{\\text{diff}}/\\sqrt{n_{\\text{diff}}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{\\sigma_1^2}{n_1} + \\frac{\\sigma_2^2}{n_2}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{s_1^2}{n_1} + \\frac{s_2^2}{n_2}}}$", correct = TRUE),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{p\\left(1 - p\\right)}{n}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{p_1\\left(1 - p_1\\right)}{n_1} + \\frac{p_2\\left(1 - p_2\\right)}{n_2}}}$"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "Which distribution does the test statistic follow for this hypothesis test?",
    answer("The normal distribution"),
    answer("The $t$-distribution with $n - 1$ degrees of freedom."),
    answer("The $t$-distribution with $\\min\\{n_1 - 1, n_2 - 1\\}$ degrees of freedom.", correct = TRUE),
    answer("The $t$-distribution with $n_{\\text{diff}} - 1$ degrees of freedom."),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

Now we'll begin building the test statistic. Use the code block below to input or compute the point estimate for our population parameter.
```{r prob2-9, exercise = TRUE}

```

```{r prob2-9-check}
grade_result(
  pass_if(~ (abs(.result) == 5))
)
```

###

Use the code block below to input or compute the null value.
```{r prob2-10, exercise = TRUE}

```

```{r prob2-10-check}
grade_result(
  pass_if(~ (.result == 0))
)
```

###

Use the code block below to input or compute the standard error.
```{r prob2-11, exercise = TRUE}

```

```{r prob2-11-check}
grade_result(
  pass_if(~ (abs(.result - sqrt((4.5^2/14) + (6.2^2/9))) < 0.01))
)
```

###

Use the code block below to input or compute the test statistic.
```{r prob2-12, exercise = TRUE}

```

```{r prob2-12-check}
grade_result(
  pass_if(~ (abs(abs(.result) - ((5 - 0)/sqrt((4.5^2/14) + (6.2^2/9)))) < 0.01))
)
```

###

Use the code block below to compute the $p$-value associated with this test.
```{r prob2-13, exercise = TRUE}

```

```{r prob2-13-check}
grade_result(
  pass_if(~ (abs(.result - (1 - pt(((5 - 0)/sqrt((4.5^2/14) + (6.2^2/9))), df = 8))) < 0.0005))
)
```

###

Answer the following to complete the hypothesis test.

```{r prob2-14, echo = FALSE}
quiz(
  question_radio(
    "What is the result of the test?",
    answer("Since $p \\geq \\alpha$, we do not have enough evidence to reject the null hypothesis."),
  answer("Since $p \\geq \\alpha$, we accept the null hypothesis."),
  answer("Since $p < \\alpha$, we reject the null hypothesis and accept the alternative hypothesis.", correct = TRUE),
  answer("Since $p < \\alpha$, we fail to reject the null hypothesis."),
  answer("It is impossible to determine."),
  allow_retry = TRUE,
  random_answer_order = TRUE
  ),
  question_radio(
  "The result of the test means that",
  answer("The sample data did not provide significant evidence to suggest that the average age of men on Death Row is younger than the average age of women on Death Row."),
  answer("The sample data proved that the average age of men on Death Row is not younger than the average age of women on Death Row."),
  answer("The sample data provided significant evidence to suggest that the average age of men on Death Row is younger than the average age of women on Death Row", correct = TRUE),
  answer("The sample data provided significant evidence to suggest that the sample average age of men on Death Row is younger than the sample average age of women on Death Row."),
  allow_retry = TRUE,
  random_answer_order = TRUE
  )
)
```

## Problem 3

**Problem 3:** A recent random sample of 30 customers who switched to a car insurance company boasting that new customers save an average of \$534 per year resulted in an average savings of \$525 with a standard deviation of \$40. Find a 90% confidence interval for the true mean savings of customers switching to this company. (In case of emergency, [see worked solution here](www/SampleInferenceCarInsurance.pdf){target="_blank"})

###

```{r prob3-1, echo = FALSE}
question_radio(
  "To answer the question as asked, we should",
  answer("Construct a confidence interval", correct = TRUE),
  answer("Conduct a hypothesis test"),
  answer("Compute a probability"),
  answer("Find a required sample size"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob3-2, echo = FALSE}
question_radio(
  "What is the desired level of confidence for this confidence interval?",
  answer("90%", correct = TRUE),
  answer("95%"),
  answer("98%"),
  answer("99%"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob3-3, echo = FALSE}
question_radio(
  "Is your confidence interval being build to capture a mean ( $\\mu$ ), a proportion ( $p$ ), or something else?",
  answer("A mean or difference in means", correct = TRUE),
  answer("A proportion or difference in proportions"),
  answer("Something else altogether"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob3-4, echo = FALSE}
question_radio(
  "Does the population parameter to be captured belong to a single group or is it a comparison of multiple groups?",
  answer("The population parameter belongs to a single group.", correct = TRUE),
  answer("The population parameter involves a comparison of two groups."),
  answer("The population parameter involves a comparison of more than two groups."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

```{r prob3-5, echo = FALSE}
question_radio(
  "Do we know the population standard deviation ( $\\sigma$ ) for dollars saved?",
  answer("Yes, the population standard deviation is known.", message = "Oops. We are presented with a standard deviation in savings but that standard deviation belongs only to the observations in the sample, not the entire population."),
  answer("No, the population standard deviation is not known.", correct = TRUE),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob3-6, echo = FALSE}
quiz(
  question_radio(
    "Which standard error formula should be used in the construction of the confidence interval?",
    answer("$\\displaystyle{S_E = \\sigma/\\sqrt{n}}$"),
    answer("$\\displaystyle{S_E = s/\\sqrt{n}}$", correct = TRUE),
    answer("$\\displaystyle{S_E = \\sigma_{\\text{diff}}/n_{\\text{diff}}}$"),
    answer("$\\displaystyle{S_E = s_{\\text{diff}}/\\sqrt{n_{\\text{diff}}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{\\sigma_1^2}{n_1} + \\frac{\\sigma_2^2}{n_2}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{s_1^2}{n_1} + \\frac{s_2^2}{n_2}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{p\\left(1 - p\\right)}{n}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{p_1\\left(1 - p_1\\right)}{n_1} + \\frac{p_2\\left(1 - p_2\\right)}{n_2}}}$"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "Which distribution should be used to identify the critical value?",
    answer("The normal distribution"),
    answer("The $t$-distribution with $n - 1$ degrees of freedom.", correct = TRUE),
    answer("The $t$-distribution with $\\min\\{n_1 - 1, n_2 - 1\\}$ degrees of freedom."),
    answer("The $t$-distribution with $n_{\\text{diff}} - 1$ degrees of freedom."),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

Use the code block below to input or compute the critical value which should be used in the construction of the confidence interval.

```{r prob3-7, exercise = TRUE}

```

```{r prob3-7-check}
grade_result(
  pass_if(~ (abs(abs(.result) - qt(0.95, df = 30 - 1)) < 0.005))
)
```

###

Use the code block below to input or compute the point estimate.

```{r prob3-8, exercise = TRUE}

```

```{r prob3-8-check}
grade_result(
  pass_if(~ (.result == 525))
)
```

###

Use the code block below to input or compute the standard error.

```{r prob3-9, exercise = TRUE}

```

```{r prob3-9-check}
grade_result(
  pass_if(~ (abs(.result - (40/sqrt(30))) < 0.01))
)
```

###

Use the code block below to compute the lower bound for our confidence interval.

```{r prob3-10, exercise = TRUE}

```

```{r prob3-10-check}
grade_result(
  pass_if(~ (abs(.result - (525 - (qt(0.95, df = 30 - 1)*40/sqrt(30)))) < 0.01))
)
```

###

Use the code block below to compute the upper bound for our confidence interval.

```{r prob3-11, exercise = TRUE}

```

```{r prob3-11-check}
grade_result(
  pass_if(~ (abs(.result - (525 + (qt(0.95, df = 30 - 1)*40/sqrt(30)))) < 0.01))
)
```

###

```{r prob3-12, echo = FALSE}
question_radio(
  "The interpretations of the values you identified are:",
  answer("We are 90% confident that the true average savings customers should expect when switching to this provider is between the lower bound and upper bounds calculated.", correct = TRUE),
  answer("90% of the savings amounts customers realize when switching to this provider will fall between the lower and upper bounds calculated."),
  answer("We are 90% confident that the average savings customers in this sample experienced when switching to this provider fall between the lower and upper bounds calculated"),
  answer("There is a 90% chance a customer switching to this provider will realize a savings between the lower and upper bounds calculated."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

## Problem 4

**Problem 4:** In January 2011, The Marist Poll published a report stating that 66% of adults nationally think licensed drivers should be required to retake their road test once they reach 65 years of age. In that same year, 200 random citizens of New Hampshire were asked whether they were in favor of an additional road test for drivers at 65 years of age, 118 responded that they were in favor. Construct a 95% confidence interval for the true proportion of New Hampshirites who are in favor of this proposal. (In case of emergency, [see worked solution here](www/SampleInferenceMaristPoll.pdf){target="_blank"} -- you're on your own from here on out though!)

###

```{r prob4-1, echo = FALSE}
question_radio(
  "To answer the question as asked, we should",
  answer("Construct a confidence interval", correct = TRUE),
  answer("Conduct a hypothesis test"),
  answer("Compute a probability"),
  answer("Find a required sample size"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob4-2, echo = FALSE}
question_radio(
  "What is the desired level of confidence for this confidence interval?",
  answer("90%"),
  answer("95%", correct = TRUE),
  answer("98%"),
  answer("99%"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob4-3, echo = FALSE}
question_radio(
  "Is your confidence interval being build to capture a mean ( $\\mu$ ), a proportion ( $p$ ), or something else?",
  answer("A mean or difference in means"),
  answer("A proportion or difference in proportions", correct = TRUE),
  answer("Something else altogether"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

```{r prob4-4, echo = FALSE}
question_radio(
  "Does the population parameter to be captured belong to a single group or is it a comparison of multiple groups?",
  answer("The population parameter belongs to a single group.", correct = TRUE),
  answer("The population parameter involves a comparison of two groups."),
  answer("The population parameter involves a comparison of more than two groups."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

```{r prob4-5, echo = FALSE}
quiz(
  question_radio(
    "Which standard error formula should be used in the construction of the confidence interval?",
    answer("$\\displaystyle{S_E = \\sigma/\\sqrt{n}}$"),
    answer("$\\displaystyle{S_E = s/\\sqrt{n}}$"),
    answer("$\\displaystyle{S_E = \\sigma_{\\text{diff}}/n_{\\text{diff}}}$"),
    answer("$\\displaystyle{S_E = s_{\\text{diff}}/\\sqrt{n_{\\text{diff}}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{\\sigma_1^2}{n_1} + \\frac{\\sigma_2^2}{n_2}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{s_1^2}{n_1} + \\frac{s_2^2}{n_2}}}$"),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{p\\left(1 - p\\right)}{n}}}$", correct = TRUE),
    answer("$\\displaystyle{S_E = \\sqrt{\\frac{p_1\\left(1 - p_1\\right)}{n_1} + \\frac{p_2\\left(1 - p_2\\right)}{n_2}}}$"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "Which distribution should be used to identify the critical value?",
    answer("The normal distribution", correct = TRUE),
    answer("The $t$-distribution with $n - 1$ degrees of freedom."),
    answer("The $t$-distribution with $\\min\\{n_1 - 1, n_2 - 1\\}$ degrees of freedom."),
    answer("The $t$-distribution with $n_{\\text{diff}} - 1$ degrees of freedom."),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

Use the code block below to input or compute the critical value which should be used in the construction of the confidence interval.

```{r prob4-6, exercise = TRUE}

```

```{r prob4-6-check}
grade_result(
  pass_if(~ (abs(abs(.result) - qnorm(0.975)) < 0.005))
)
```

###

Use the code block below to input or compute the point estimate.

```{r prob4-7, exercise = TRUE}

```

```{r prob4-7-check}
grade_result(
  pass_if(~ abs(.result - (118/200)) < 0.0005)
)
```

###

Use the code block below to input or compute the standard error.

```{r prob4-8, exercise = TRUE}

```

```{r prob4-8-check}
grade_result(
  pass_if(~ (abs(.result - (sqrt(((118/200)*(1 - (118/200)))/200))) < 0.0005))
)
```

###

Use the code block below to compute the lower bound for our confidence interval.

```{r prob4-9, exercise = TRUE}

```

```{r prob4-9-check}
grade_result(
  pass_if(~ (abs(.result - ((118/200) - (qnorm(0.975)*sqrt(((118/200)*(1 - (118/200)))/200)))) < 0.0005))
)
```

###

Use the code block below to compute the upper bound for our confidence interval.

```{r prob4-10, exercise = TRUE}

```

```{r prob4-10-check}
grade_result(
  pass_if(~ (abs(.result - ((118/200) + (qnorm(0.975)*sqrt(((118/200)*(1 - (118/200)))/200)))) < 0.0005))
)
```

###

```{r prob4-11, echo = FALSE}
question_radio(
  "The interpretations of the values you identified are:",
  answer("We are 95% confident that the true proportion of New Hampshirites favoring the policy is between the lower bound and upper bounds calculated.", correct = TRUE),
  answer("95% of New Hampshirites favor the policy."),
  answer("We are 95% confident that the sample proportion of New Hampshirites favoring the policy fall between the lower and upper bounds calculated"),
  answer("There is a 95% chance true proportion of New Hampshirites favoring the policy falls between the lower and upper bounds calculated."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

## Submit

```{r context="server"}
learnrhash::encoder_logic(strip_output = TRUE)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = shiny::div(
    "If your instructor is asking you to submit evidence of your completed notebook using a hash code, you may generate that code below. Use the clipboard icon to ensure that you are copying the entire hash."
    )
  )
```
