---
title: "Topic 6: The Normal Distribution"
tutorial:
  id: "NormalDist.Topic6.AppliedStats"
  version: 1.0
output: 
  learnr::tutorial:
    progressive: TRUE
    allow_skip: TRUE
    
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)
#library(tidyverse)
library(dplyr)
library(ggplot2)
library(learnr)
library(gradethis)


tutorial_options(exercise.timelimit = 60, exercise.checker = gradethis::grade_learnr)

#set.seed(round(as.numeric(Sys.time()), -2))
set.seed(241)

boundary1 <- round(runif(1, 0.75, 2), 2)
boundary2 <- round(runif(1, -2, -1.5), 2)
boundary3 <- round(runif(1, -2.5, -0.75), 2)
boundary4 <- round(runif(1, 0.75, 2.5), 2)
boundary5a <- round(runif(1, -1.5, -0.75), 2)
boundary5b <- round(runif(1, 1, 2.5), 2)
boundary6a <- round(runif(1, -1.5, -0.75), 2)
boundary6b <- round(runif(1, 1.5, 2.5), 2)
boundary7 <- round(runif(1, 1, 2.5), 2)

chickPoxSampSize <- sample(50:150, size = 1)
#knitr::opts_chunk$set(eval = FALSE)
```



## Topic 6: The Normal Distribution

###

Throughout this workbook we'll investigate the probability distribution that is most central to our study of statistics: the normal distribution.  If we are confident that our data are nearly normal, that opens the door to many powerful statistical methods. This workbook gives you practice in working with normally distributed data.

## Objectives

**Workbook Objectives:** After completing this workbook you should be able to:

+ Compute probabilities of events well-modeled by a normal distribution.
+ Given a variable $X$ which follows an assumed normal distribution, compute and interpret various percentile thresholds for $X$.
+ Identify scenarios to which the normal or binomial distributions can be applied, and use them to answer various probability-related questions.

###

## The Normal Distribution

**Definition:** If a random variable $X$ is normally distributed with mean $\mu$ and standard deviation $\sigma$, we often write $X\sim N\left(\mu, \sigma\right)$. Three different normal distributions appear below.

+ In **<span style = "color:blue">blue</span>** is a normal distribution with $\mu = 0$ and $\sigma = 5$
+ In **<span style = "color:red">red</span>** is a normal distribution with $\mu = 0$ and $\sigma = 0.5$
+ In **black** is a normal distribution with $\mu = 0$ and $\sigma = 1$ (the so-called *Standard Normal Distribution*)

```{r three-normals, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE}
ggplot() + geom_line(mapping = aes(x = seq(-15, 15, length.out = 200), y = dnorm(seq(-15, 15, length.out = 200), 0, 5)), color = "blue", lwd = 1.25) + geom_line(mapping = aes(x = seq(-15, 15, length.out = 200), y = dnorm(seq(-15, 15, length.out = 200), 0, 0.5)), color = "red", lwd = 1.25) + geom_line(mapping = aes(x = seq(-15, 15, length.out = 200), y = dnorm(seq(-15, 15, length.out = 200), 0, 1)), color = "black", lwd = 1.25) + labs(title = "Three Normal Distributions", x = "", y= "")
```

###

Notice that all three distributions are bell-shaped and are centered at their mean ($\mu = 0$). The larger the standard deviation, the shorter and wider the curve, while the smaller the standard deviation, the taller and more narrow the curve.

###

Given that $X\sim N\left(\mu, \sigma\right)$, we can compute probabilities associated with observed values of $X$ by finding the corresponding area beneath the normal curve with mean $\mu$ and standard deviation $\sigma$.

###

**Properties of the Normal Distribution:** We have the following properties associated with the normal distribution. Consider $X\sim N\left(\mu, \sigma\right)$.

+ The area beneath the entire distribution is 1 (since this is equivalent to the probability that $X$ takes on any of its possible values).

<center>
```{r shade-norm, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(-3, 3), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("The shaded area is 1") + scale_x_continuous(breaks = c(0), labels = c(expression(mu)))  + theme(axis.text.x = element_text(size = 14))
```
</center>

###

+ $\displaystyle{\mathbb{P}\left[X\leq \mu\right] = \mathbb{P}\left[X\geq \mu\right] = 0.5}$ (the area underneath a full half of the distribution is 0.5)

<center>
```{r shade-norm-half, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(-3, 0), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("The shaded area is 0.5") + scale_x_continuous(breaks = c(0), labels = c(expression(mu))) + theme(axis.text.x = element_text(size = 14))

ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(0, 3), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("The shaded area is also 0.5") + scale_x_continuous(breaks = c(0), labels = c(expression(mu))) + theme(axis.text.x = element_text(size = 14))
```
</center>

###

+ The distribution is symmetric. In symbols, $\mathbb{P}\left[X\leq \mu - k\right] = \mathbb{P}\left[X \geq \mu + k\right]$ for any $k$.

<center>
```{r shade-norm-symmetric, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(-3, -1.25), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("A shaded area in the left \ntail") + scale_x_continuous(breaks = c(0, -1.25), labels = c(expression(mu), expression(mu - k))) + theme(axis.text.x = element_text(size = 14))

ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(1.25, 3), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("The same shaded area in the \nright tail") + scale_x_continuous(breaks = c(0, 1.25), labels = c(expression(mu), expression(mu + k))) + theme(axis.text.x = element_text(size = 14))
```
</center>

###

+ $\displaystyle{\mathbb{P}\left[X = k\right] = 0}$ (the probability that $X$ takes on any prescribed value exactly is $0$)

<center>
```{r shade-norm-exact, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(0.98, 1.02), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("The area is 0") + ylim(c(0, 0.42)) + scale_x_continuous(breaks = c(0, 1), labels = c(expression(mu), "k")) + theme(axis.text.x = element_text(size = 14))
```
</center>

###

Sometimes it is useful to be able to estimate probabilities or to estimate the proportion of a population that falls into a range as long as the population is nearly normal. A convenient rule of thumb is the *Empirical Rule*.

**The Empirical Rule:** If $X\sim N\left(\mu, \sigma\right)$, then

+ $\mathbb{P}\left[\mu - \sigma \leq X\leq \mu + \sigma\right] \approx 0.67$ -- that is, about 67% of observations lie within one standard deviation of the mean.
+ $\mathbb{P}\left[\mu - 2\sigma \leq X\leq \mu + 2\sigma\right] \approx 0.95$ -- that is, about 95% of observations lie within two standard deviations of the mean.
+ $\mathbb{P}\left[\mu - 3\sigma \leq X\leq \mu + 3\sigma\right] \approx 0.997$ -- that is, about 99.7% of observations lie within three standard deviations of the mean.

###

For each of the following, assume that $X\sim N\left(\mu = 85, \sigma = 5\right)$

```{r empirical-rule-questions, echo = FALSE}
quiz(
  question_radio(
    "Use the *Empirical Rule* to approximate $\\mathbb{P}\\left[80\\leq X\\leq 90\\right]$",
    answer("50%"),
    answer("67%", correct = TRUE),
    answer("95%"),
    answer("Nearly 100%"),
    answer("It is impossible to tell"),
    allow_retry = TRUE
  ),
  question_checkbox(
    "According to the *Empirical Rule* which of the following are boundary values for which we expect about 95% of observed values of X will fall between?",
    answer("0"),
    answer("50"),
    answer("75", correct = TRUE),
    answer("80"),
    answer("90"),
    answer("95", correct = TRUE),
    answer("100"),
    allow_retry = TRUE
  )
)
```

## Standardization and $z$-scores

**Scenario:** Two students, Bob and Sally, are trying to compare how well they did on a college entrance exam. The difficulty comes in that Bob took the SAT which is known to follow an approximate normal distribution with a mean score of 1068 and a standard deviation of 210 while Sally took the ACT which also follows an approximately normal distribution but with a mean score of 20.8 and a standard deviation of 5.8. If Bob scored a 1400 on the SAT and Sally scored a 31 on the ACT, who scored relatively higher?

How do we answer this question? We'll see two methods.

### 

**Method 1:** We can standardize the test scores so that they have comparable units. 

+ **Definition:** If an observation $x$ comes from a nearly normal population with mean $\mu$ and standard deviation $\sigma$ then we compute $z$-score associated with $x$ as follows: 

$$\displaystyle{z = \frac{x - \mu}{\sigma}}$$

An observation's $z$-score is simply the number of standard deviations it falls above or below the mean. 

###

Use the code block below to compute Bob and Sally's $z$-scores and answer the questions that follow.

```{r compute-z-scores, exercise = TRUE}

```

```{r SAT-ACT-z-scores, echo = FALSE}
quiz(
  question_radio(
    "Which of the following is the $z$-score corresponding to Bob? (round to two decimal places)",
    answer("1.58", correct = TRUE),
    answer("1394.91"),
    answer("-1.58"),
    answer("1.76"),
    answer("-1.76"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "Which of the following is the $z$-score corresponding to Sally? (round to two decimal places)",
    answer("1.58"),
    answer("27.41"),
    answer("-1.58"),
    answer("1.76", correct = TRUE),
    answer("-1.76"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "Who did relatively better on their standardized exam, Bob or Sally? Why?",
    answer("Sally, since her score is 1.76 standard deviations above the mean, while Bob's was only 1.58 standard deviations above the mean.", correct = TRUE),
    answer("Bob, since his $z$-score was 1.58 which is closer to the mean than Sally's $z$-score of 1.76."),
    answer("Bob, since his score was 1400 and Sally's score was only 31."),
    answer("Bob, since the SAT had a larger standard deviation it is a harder exam."),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

**A recap on $z$-scores:** We can use $z$-scores as a common unit for comparing observations from completely different populations (such as SAT scores and ACT scores). Here's a recap of the most important information so far:

+ If an observation $x$ comes from a nearly normal population with mean $\mu$ and standard deviation $\sigma$, we can compute it's $z$-score using the formula: $\displaystyle{z = \frac{x - \mu}{\sigma}}$.
+ A $z$-score measures the number of standard deviations which an observation falls above or below the mean.

  + A positive $z$-score means that an observation was above the mean.
  + A negative $z$-score means that an observation was below the mean.
  + The larger a $z$-score is in absolute value, the further the corresponding observation falls from the mean. That is, the larger the magnitude of a $z$-score, the further into the tail of the distribution the corresponding observation falls.
  
###

**Method 2:** We can compute the *percentile* corresponding to Bob's SAT score and the *percentile* corresponding to Sally's ACT score.
  
  + **Definition:** Given an observation $x$ from a population -- the *percentile* corresponding to $x$ is the proportion of the population which falls below $x$.
  
###

<div class = "row">
<div class = "col-md-6">
Bob's percentile corresponds to the shaded area in the distribution below.

```{r bob-SAT-percentile, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(500, 1600))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 1068, sd = 210), xlim = c(500, 1400), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(500, 1600, length.out = 200), y = dnorm(seq(500, 1600, length.out = 200), 1068, 210)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("SAT Distribution") + xlim(c(500, 1600)) + scale_x_continuous(breaks = c(1068, 1400), labels = c(expression(mu = 1068), "Bob")) + theme(axis.text.x = element_text(size = 14))
```
</div>
<div class = "col-md-6">
Sally's percentile corresponds to the shaded area in the distribution below.

```{r sally-SAT-percentile, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(0, 36))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 20.8, sd = 5.8), xlim = c(0, 31), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(0, 36, length.out = 200), y = dnorm(seq(0, 36, length.out = 200), 20.8, 5.8)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("ACT Distribution") + xlim(c(0, 36)) + scale_x_continuous(breaks = c(20.8, 31), labels = c(expression(mu = 20.8), "Sally")) + theme(axis.text.x = element_text(size = 14))
```
</div>
</div>

### 

There are many ways to compute percentiles. Before the widespread availability of statistical software, people converted observed values to $z$-scores and then looked up the percentile in a table. Luckily R provides nice functionality for computing percentiles.

**Computing Percentiles in R:** If $X\sim N\left(\mu, \sigma\right)$, then 
$$\mathbb{P}\left[X\leq q\right] \approx \tt{pnorm(q, mean = \mu, sd = \sigma)}$$

###

The block below is preset to compute the Bob's percentile. Execute the code cell and then adapt the code to find Sally's percentile. Use your results to answer the questions below.

```{r compute-percentiles, exercise = TRUE}
pnorm(1400, 1068, 210)
```

```{r SAT-ACT-percentiles, echo = FALSE}
quiz(
  question_radio(
    "Which of the following is the percentile corresponding to Bob? (round to four decimal places)",
    answer("0.9431", correct = TRUE),
    answer("0.0569"),
    answer("0.4431"),
    answer("0.9608"),
    answer("0.0392"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "Which of the following is the percentile corresponding to Sally? (round to two decimal places)",
    answer("0.9431"),
    answer("0.0569"),
    answer("0.4431"),
    answer("0.9607", correct = TRUE),
    answer("0.0392"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "Who did relatively better on their standardized exam, Bob or Sally? Why?",
    answer("Sally, since she scored in a higher percentile than Bob", correct = TRUE),
    answer("Bob, since he scored in a lower percentile than Sally"),
    answer("Bob, since he scored closer to the mean"),
    answer("Bob, since his percentile score was so high even though the standard deviation on the SAT was so large."),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

We'll make good use of this second method for a while, but don't forget about standardization and $z$-scores. We'll need that strategy quite often later in our course! For now, let's move on to practicing with finding probabilities from a normal distribution using R's `pnorm()` function.

## Computing probability from a normal distribution

Through this section you'll be getting practice finding probabilities by using R's `pnorm()` function to compute areas. Remember that the `pnorm()` function takes three arguments -- the first is a $\tt{boundary}$ value, the second is the $\tt{mean}$ of the distribution, and the third is the $\tt{standard~deviation}$. The value returned by `pnorm()` is the area to the <u>left</u> of the provided boundary value in the distribution with the mean and standard deviation you provided.

###

For these first few questions I'll draw pictures for you, but you should be prepared to draw your own shortly.

###

<div class = "row">
<div class = "col-md-6">

**Question 1:** Use the code block below to find $\mathbb{P}\left[Z < \right.$ `r boundary1` $\left.\right]$ -- Remember that $Z\sim N\left(\mu = 0, \sigma = 1\right)$.

```{r pnorm-ex1-left-setup}

```

```{r pnorm-ex1-left, exercise = TRUE}

```

```{r pnorm-ex1-left-check}
grade_result(
  pass_if(~ (abs(.result - pnorm(boundary1, 0, 1)) < 0.001))
)
```
</div>
<div class = "col-md-5">
```{r pnorm-ex1-left-plot, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(-3, boundary1), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("") + xlim(c(-3, 3)) + scale_x_continuous(breaks = c(0, boundary1)) + theme(axis.text.x = element_text(size = 12, angle = 45))
```
</div>
</div>

###

<div class = "row">
<div class = "col-md-6">

**Question 2:** Find $\mathbb{P}\left[Z < \right.$ `r boundary2` $\left.\right]$.

```{r pnorm-ex2-left, exercise = TRUE}

```

```{r pnorm-ex2-left-check}
grade_result(
  pass_if(~ (abs(.result - pnorm(boundary2, 0, 1)) < 0.001))
)
```
</div>
<div class = "col-md-5">
```{r pnorm-ex2-left-plot, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(-3, boundary2), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("") + xlim(c(-3, 3)) + scale_x_continuous(breaks = c(0, boundary2)) + theme(axis.text.x = element_text(size = 12, angle = 45))
```
</div>
</div>

###

<div class = "row">
<div class = "col-md-6">

**Question 3:** Find $\mathbb{P}\left[Z > \right.$ `r boundary3` $\left.\right]$.

```{r pnorm-ex3-left, exercise = TRUE}

```

```{r pnorm-ex3-left-check}
grade_result(
  pass_if(~ (abs(.result - (1 - pnorm(boundary3, 0, 1))) < 0.001))
)
```
</div>
<div class = "col-md-5">
```{r pnorm-ex3-left-plot, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(boundary3, 3), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("") + xlim(c(-3, 3)) + scale_x_continuous(breaks = c(0, boundary3)) + theme(axis.text.x = element_text(size = 12, angle = 45))
```
</div>
</div>

###

<div class = "row">
<div class = "col-md-6">

**Question 4:** Find $\mathbb{P}\left[Z > \right.$ `r boundary4` $\left.\right]$.

```{r pnorm-ex4-left, exercise = TRUE}

```

```{r pnorm-ex4-left-check}
grade_result(
  pass_if(~ (abs(.result - (1 - pnorm(boundary4, 0, 1))) < 0.001))
)
```
</div>
<div class = "col-md-5">
```{r pnorm-ex4-left-plot, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(boundary4, 3), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("") + xlim(c(-3, 3)) + scale_x_continuous(breaks = c(0, boundary4)) + theme(axis.text.x = element_text(size = 12, angle = 45))
```
</div>
</div>

###

<div class = "row">
<div class = "col-md-6">

**Question 5:** Find $\mathbb{P}\left[\right.$ `r boundary5a` $< Z <$  `r boundary5b` $\left.\right]$.

```{r pnorm-ex5-left, exercise = TRUE}

```

```{r pnorm-ex5-left-check}
grade_result(
  pass_if(~ (abs(.result - (pnorm(boundary5b, 0, 1) - pnorm(boundary5a, 0, 1))) < 0.001))
)
```
</div>
<div class = "col-md-5">
```{r pnorm-ex5-left-plot, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(boundary5a, boundary5b), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("") + xlim(c(-3, 3)) + scale_x_continuous(breaks = c(boundary5a, 0, boundary5b)) + theme(axis.text.x = element_text(size = 12, angle = 45))
```
</div>
</div>

###

<div class = "row">
<div class = "col-md-6">

**Question 6:** Find $\mathbb{P}\left[Z < \right.$ `r boundary6a` $\text{ or } Z >$ `r boundary6b` $\left.\right]$.

```{r pnorm-ex6-left, exercise = TRUE}

```

```{r pnorm-ex6-left-check}
grade_result(
  pass_if(~ (abs(.result - (pnorm(boundary6a, 0, 1) +(1 - pnorm(boundary6b, 0, 1)))) < 0.001))
)
```
</div>
<div class = "col-md-5">
```{r pnorm-ex6-left-plot, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(-3, boundary6a), fill = "steelblue", alpha = 0.5) + geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(boundary6b, 3), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("") + xlim(c(-3, 3)) + scale_x_continuous(breaks = c(boundary6a, 0, boundary6b)) + theme(axis.text.x = element_text(size = 12, angle = 45))
```
</div>
</div>

###

<div class = "row">
<div class = "col-md-6">

**Question 7:** Find $\mathbb{P}\left[\left|Z\right| > \right.$ `r boundary7` $\left.\right]$.

```{r pnorm-ex7-left, exercise = TRUE}

```

```{r pnorm-ex7-left-check}
grade_result(
  pass_if(~ (abs(.result - (pnorm(-boundary7, 0, 1) +(1 - pnorm(boundary7, 0, 1)))) < 0.001))
)
```
</div>
<div class = "col-md-5">
```{r pnorm-ex7-left-plot, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(-3, -boundary7), fill = "steelblue", alpha = 0.5) + geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(boundary7, 3), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("") + xlim(c(-3, 3)) + scale_x_continuous(breaks = c(-boundary7, 0, boundary7)) + theme(axis.text.x = element_text(size = 12, angle = 45))
```
</div>
</div>

###

Through the last seven problems you only worked with the standard normal distribution -- that's the $Z$-distribution, which is $N\left(\mu = 0, \sigma = 1\right)$. We can find probabilities from arbitrary normal distributions (normal distributions with any mean and any standard deviation) using R's `pnorm()` functionality -- just supply the appropriate `mean` and `sd` arguments to `pnorm()` instead of the 0 and 1 that we passed earlier.

## Finding percentile cutoffs on a normal distribution

<div class = "row">
<div class = "col-md-7">
Recall from earlier that the $p^{th}$ percentile of a random variable $X$ is the value $x^*$ such that $\mathbb{P}\left[X < x^*\right] = p$.
</div>
<div class = "col-md-5">
```{r percentile-cutoff-def, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, results = FALSE, fig.width = 3, fig.height = 3}
ggplot(NULL, aes(c(-3, 3))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 0, sd = 1), xlim = c(-3, 1.96), fill = "steelblue", alpha = 0.5) + geom_line(mapping = aes(x = seq(-3, 3, length.out = 200), y = dnorm(seq(-3, 3, length.out = 200), 0, 1)), color = "blue", lwd = 1.25) + xlab("") + ylab("") + ggtitle("") + xlim(c(-3, 3)) + scale_x_continuous(breaks = c(0, 1.96), labels = c(expression(mu), "x*")) + theme(axis.text.x = element_text(size = 12))
```
</div>
</div>

###

If $X\sim N\left(\mu, \sigma\right)$, then to find the cutoff $x^*$ for which $\mathbb{P}\left[X < x^*\right] = p$, we can use R's `qnorm()` function. Similar to `pnorm()`, this function takes three arguments. The first is the $\tt{area~to~the~\underline{LEFT}}$ of the desired cutoff, the second is the $\tt{mean}$ of the distribution, and the third is the $\tt{standard~deviation}$ of the distribution.

###

Recall from earlier that SAT scores followed $N\left(\mu = 1068, \sigma = 210\right)$ and ACT scores followed $N\left(\mu = 20.8, \sigma = 5.8\right)$. The code block below is set up to find the minimum required SAT score to fall in the 95th percentile (to do better than 95% of other test-takers). Execute the code and note the required score. Adapt the code to find the minimum ACT score required to fall into the top 10% of all ACT test takers. Does your answer seem right? How can you judge?

```{r find-percentile, exercise = TRUE, exercise.eval = FALSE}
qnorm(0.95, 1068, 210)
```

```{r find-percentile-check}
grade_result(
  pass_if(~ (abs(.result - qnorm(0.9, 20.8, 5.8)) < 0.1))
)
```

## Practice with the normal and binomial distributions

Through this last section you'll work through a set of problems, some of which use the normal distribution while others use the binomial distribution. It is up to you to determine which distribution should be applied in each problem. Below are a few helpful reminders:

+ The binomial distribution can be applied to scenarios of repeated trials, where each trial has two possible outcomes, and the probability of "success" on each trial remains constant (that is, trials are independent of one another). If we let $X$ be the number of successes observed in a binomial experiment with $n$ trials and a probability of success equal to $p$, then we have the following:
  
  + $\mathbb{P}\left[X = k\right] = \tt{dbinom(k, n, p)}$
  + $\mathbb{P}\left[X \leq k\right] = \tt{pbinom(k, n, p)}$

+ The normal distribution can be applied to scenarios where we know data follows at least a nearly-normal distribution. If $X\sim N\left(\mu, \sigma\right)$, then

  + $\mathbb{P}\left[X \leq k\right] = \tt{pnorm(k, mean = \mu, sd = \sigma)}$
  + If we wish to find the $p^{th}$ percentile of $X$ -- that is, we want to find the $\tt{boundary}$ value for which the probability of lying to the left is $p$, then we utilize $\tt{boundary = qnorm(p, mean = \mu, sd = \sigma)}$
  
###

**Practice Problem 1:** The National Vaccine Information Center estimates that 90% of Americans have had chickenpox by the time they reach adulthood. Suppose we take a random sample of `r chickPoxSampSize` American adults. Answer each of the following:

+ Compute the expected number of adults in our sample who will have had chickenpox.
```{r chick-pox-exp, exercise = TRUE}

```

```{r chick-pox-exp-check}
grade_result(
  pass_if(~ (abs(.result - (chickPoxSampSize*0.9)) < 0.1))
)
```

+ Compute the standard deviation in number of adults in samples of size `r chickPoxSampSize` who will have had chickenpox.
```{r chick-pox-sd, exercise = TRUE}

```

```{r chick-pox-sd-check}
grade_result(
  pass_if(~ (abs(.result - (sqrt(chickPoxSampSize*0.9*0.1))) < 0.1))
)
```

```{r surprised-if, echo = FALSE}
question_checkbox(
  paste0("Would you be surprised if your sample contained at most ", floor((chickPoxSampSize*0.9) - (3.5*sqrt(chickPoxSampSize*0.9*0.1))), " adults who have had chicken pox? Why? (select all that are appropriate)"),
  answer("Yes. This result would fall more than two standard deviations away from the expected value.", correct = TRUE),
  answer("Yes. If our assumptions are correct, the probability of this occurring would be less than 5%.", correct = TRUE),
  answer("Yes. We should be surprised by any result at all -- we can't know what will happen ahead of time."),
  answer("No. A result like this one is certainly possible."),
  answer("No. This result does not fall more than two standard deviations from the expected value."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

+ Find the probability that exactly `r floor(0.9*chickPoxSampSize)` adults in your sample of size `r chickPoxSampSize` will have had chickenpox.
```{r chick-pox-prob-exact1, exercise = TRUE}

```

```{r chick-pox-prob-exact1-check}
grade_result(
  pass_if(~ (abs(.result - (dbinom(floor(0.9*chickPoxSampSize), chickPoxSampSize, 0.9))) < 0.001))
)
```

+ Find the probability that exactly `r floor(0.1*chickPoxSampSize)` adults in your sample of size `r chickPoxSampSize` will *not* have had chickenpox.
```{r chick-pox-prob-exact2, exercise = TRUE}

```

```{r chick-pox-prob-exact2-check}
grade_result(
  pass_if(~ (abs(.result - (dbinom(floor(0.1*chickPoxSampSize), chickPoxSampSize, 0.1))) < 0.001))
)
```

+ Find the probability that at most `r floor(0.75*chickPoxSampSize)` adults in your sample of size `r chickPoxSampSize` will have had chicken pox.
```{r chick-pox-prob-atmost, exercise = TRUE}

```

```{r chick-pox-prob-atmost-check}
grade_result(
  pass_if(~ (abs(.result - (pbinom(floor(0.75*chickPoxSampSize), chickPoxSampSize, 0.9))) < 0.00005))
)
```

+ Find the probability that at least `r floor(0.95*chickPoxSampSize)` adults in your sample of size `r chickPoxSampSize` will have had chicken pox.
```{r chick-pox-prob-atleast, exercise = TRUE}

```

```{r chick-pox-prob-atleast-check}
grade_result(
  pass_if(~ (abs(.result - ( 1 - pbinom((floor(0.95*chickPoxSampSize) - 1), chickPoxSampSize, 0.9))) < 0.0001))
)
```

+ Find the probability that more than `r floor(0.8*chickPoxSampSize)` but less than `r floor(0.9*chickPoxSampSize)` adults in your sample of size `r chickPoxSampSize` will have had chicken pox.
```{r chick-pox-prob-between, exercise = TRUE}

```

```{r chick-pox-prob-between-check}
grade_result(
  pass_if(~ (abs(.result - ( pbinom((floor(0.9*chickPoxSampSize) - 1), chickPoxSampSize, 0.9) - pbinom(floor(0.8*chickPoxSampSize), chickPoxSampSize, 0.9))) < 0.0001))
)
```

**Practice Problem 2:** Sophia took the Graduate Record Examination (GRE) and scored 160 on the Verbal Reasoning section and 157 on the Quantitative Reasoning section. The mean score on the Verbal Reasoning section for all test takers was 151 with a standard deviation of 7, and the mean score for the Quantitative Reasoning section for all test takers was 153 with a standard deviation of 7.67. Suppose we can assume that both score distributions are nearly normal.

+ Use the code block below to compute Sophia's $z$-score on the Quantitative Reasoning exam.
```{r z-score-QR, exercise = TRUE}

```

```{r z-score-QR-check}
grade_result(
  pass_if(~ (abs(.result - ((157 - 153)/7.67)) < 0.01))
)
```

+ Use the code block below to compute Sophia's $z$-score on the Verbal Reasoning exam.
```{r z-score-VR, exercise = TRUE}

```

```{r z-score-VR-check}
grade_result(
  pass_if(~ (abs(.result - ((160 - 151)/7)) < 0.01))
)
```

```{r GRE-compare-zscore, echo = FALSE}
question_radio(
  "What do these z-scores tell you?",
  answer("Sophia scored relatively higher on the Verbal Reasoning exam since her  Verbal Reasoning z-score was larger.", correct = TRUE),
  answer("Sophia scored relatively higher on the Quantitative Reasoning exam since her Quantitative Reasoning z-score was larger."),
  answer("Sophia scored relatively higher on the Quantitative Reasoning exam since her Quantitative Reasoning z-score was closer to 0."),
  answer("It is impossible to tell which exam Sophia performed better on since the means and standard deviations of the score distributions are different."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

+ Find the proportion of test takers Sophia scored higher than on the Quantitative Reasoning exam (that is, find her percentile).
```{r GRE-QR-percentile, exercise = TRUE}

```

```{r GRE-QR-percentile-check}
grade_result(
  pass_if(~ (abs(.result - pnorm(157, 153, 7.67)) < 0.001))
)
```

+ Find the proportion of test takers doing *better than* Sophia on the Verbal Reasoning exam.
```{r GRE-VR-percent-above, exercise = TRUE}

```

```{r GRE-VR-percent-above-check}
grade_result(
  pass_if(~ (abs(.result - (1 - pnorm(160, 151, 7))) < 0.001))
)
```

## Submit

```{r context="server"}
learnrhash::encoder_logic(strip_output = TRUE)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = shiny::div(
    "If your instructor is asking you to submit evidence of your completed notebook using a hash code, you may generate that code below. Use the clipboard icon to ensure that you are copying the entire hash."
    )
  )
```

## Summary

Nice job getting through this workbook. Your hard work here will pay off as we move through much of the remainder of our course. Here are the major points we touched on.

+ A normal distribution is approximately bell-shaped and can be described by its mean $\mu$ and standard deviation $\sigma$. 
+ As a shorthand, we often write $N\left(\mu, \sigma\right)$ to mean the normal distribution with mean $\mu$ and standard deviation $\sigma$.
+ The *Empirical Rule* is a "rule of thumb" that states that, if data is normally distributed we expect:

  + Approximately 67% of observations fall within one standard deviation of the mean.
  + Approximately 95% of observations fall within two standard deviations of the mean.
  + Approximately 99.7% of observations fall within three standard deviations of the mean.

+ We can interpret areas underneath the normal distribution to be probabilities.
+ If $X\sim N\left(\mu, \sigma\right)$, then $\mathbb{P}\left[X\leq k\right] = \tt{pnorm(k, mean = \mu, sd = \sigma)}$

  + Remember that $\tt{pnorm()}$ always returns the area to the <u>left</u> of the boundary value $k$.
+ If $X\sim N\left(\mu, \sigma\right)$, then the $p^{th}$ percentile of $X$ (the cutoff for which the proportion of the population falling below is $p$), is given by $\tt{qnorm(p, mean = \mu, sd = \sigma)}$

  + Remember that $p$ must be the area to the <u>left</u> of the desired cutoff value.
