---
title: "Chapter 1: An Introduction to Data"
tutorial:
  id: "IntroData.SNHU.ISRS"
  version: 1.0
output: 
  learnr::tutorial:
    progressive: TRUE
    
runtime: shiny_prerendered
---

```{r global-options, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)
library(MASS)
library(tidyverse)
library(learnr)
library(gradethis)
library(openintro)

houses <- read.csv("https://raw.githubusercontent.com/NikhilKumarMutyala/Linear-Regression-from-scartch-on-KC-House-Dataset/master/kc_house_data.csv")
houses$exterior <- sample(c("brick", "vinyl siding", "shingles", "NA"), size = nrow(houses), prob = c(0.15, 0.7, 0.12, 0.03), replace = TRUE)
houses$month <- sample(1:12, size = nrow(houses), replace = TRUE)
housesSubset <- houses[ , c(1, 23, 3, 6, 7, 9, 22)]

knitr::opts_chunk$set(exercise.checker = gradethis::grade_learnr)
```

# An Introduction to Data

Throughout Chapter 1 of our textbook the authors reference the `email50` dataset. You'll have an opportunity to explore a new dataset concerning real estate from King County, WA throughout this interactive document. The `homes` data is shown below in a convenient form where the rows of data represent *records* (sometimes called **observations**), and the columns represent **variables** (sometimes called **features**). We'll see later in our course that such data is called *tidy*.

```{r showData, exercise = FALSE}
head(housesSubset[sample(1:nrow(housesSubset), size = nrow(housesSubset), replace = FALSE), ])
```

You can see the first six rows of the dataset above. A simplified *data dictionary* (a map of column names and explanations) appears below.

+ `id` provides the property identification number.
+ `month` gives the month that the property was listed for sale.
+ `price` is the listing price of the property in US dollars (\$).
+ `sqft_living` gives the finished square footage of the home.
+ `sqft_lot` gives the square footage of the property (land).
+ `waterfront` indicates whether the property is waterfront.
+ `exterior` provides a description of the exterior covering of the home.

## Variable Types

Variables which take on numerical values <u>and</u> for which measures such as the mean, median, or standard deviation are meaningful are referred to as **numerical**. Non-numerical variables which serve to group data into categories are called **categorical** (eg. the color of a car or an area code prefix for a telephone number). It is possible for data to be neither numerical nor categorical -- for example, a **unique identifier** is a non-numeric variable which is unique to each record. A timestamp may be an example of a unique identifier -- all is not lost, however, with a bit of pre-processing we sometimes extract useful information from these columns.

Answer the following using your knowledge of the dataset and variable types.

```{r id-vars-numerical, echo = FALSE}
quiz(
question_checkbox(
"Check off each of the variables from the email dataset which are numerical",
answer("id", message = "are we able to take a meaningful average of these values?"),
answer("month", message = "is an average meaningful here?"),
answer("price", correct = TRUE),
answer("sqft_living", correct = TRUE),
answer("sqft_lot", correct = TRUE),
answer("waterfront"),
answer("exterior"),
allow_retry = TRUE,
try_again = "Be sure to select all of them! There are three."
),

question_checkbox(
"Check off each of the variables from the email dataset which are categorical",
answer("id", message = "are we able to take a meaningful average of these values?"),
answer("month", correct = TRUE),
answer("price"),
answer("sqft_living"),
answer("sqft_lot"),
answer("waterfront", correct = TRUE),
answer("exterior", correct = TRUE),
allow_retry = TRUE,
try_again = "Be sure to select all of them!"
),

question_checkbox(
"Check off each of the variables from the email dataset which are unique identifiers.",
answer("id", correct = TRUE),
answer("month"),
answer("price"),
answer("sqft_living"),
answer("sqft_lot"),
answer("waterfront"),
answer("exterior"),
allow_retry = TRUE,
try_again = "There's just one here!"
)

)
```

The **levels** of a variable are the different (unique) values that the variable takes on. For example, a dataset on student might include a variable called `ClassYear` with the levels *Freshmen*, *Sophomore*, *Junion*, *Senior*. Numerical variables also have levels -- usually there are lots of levels corresponding to a numerical variable, but if there are too few, we may be better off considering the corresponding variable to be categorical (imagine a dataset which includes a column for `Year`, but only the years 2008 and 2017 are included).

## Working with numerical data

Getting back to the `homes` dataset, let's consider the `sqft_living` variable. 

## Working with categorical data

