---
title: "Topic 7: Discrete Distributions Lab"
tutorial:
  id: "DiscreteDistributionsLab.Topic7.AppliedStats"
  version: 1.0
output: 
  learnr::tutorial:
    progressive: TRUE
    
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)
#library(tidyverse)
library(ggplot2)
library(dplyr)
library(learnr)
library(gradethis)

download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")

kobe_streak <- calc_streak(kobe$basket)

tutorial_options(exercise.timelimit = 60, exercise.checker = gradethis::grade_learnr)

#knitr::opts_chunk$set(eval = FALSE)
```

# Topic 7: Discrete Distributions and Simulation

## Introduction

The previous two workbooks discussed probability distributions. In this seventh workbook you'll engage in computer simulation to test a theory by building and comparing distributions. Computer simulations can be really helpful because they allow us to repeat tasks very quickly (much more quickly than we would be able to do so in real life) and they also allow us to model tasks which may be impossible or unethical to perform in real life. Simulation is a truly powerful tool for helping us understand processes in our world.

###

## Kobe Bryant and the *Hot Hand*

Basketball players who make several baskets in succession are often described as having a *hot hand*. Fans and players have long believed in the *hot hand* phenomenon, which refutes the assumption that each shot is independent of the next. However, a 1985 paper by Gilovich, Vallone, and Tversky collected evidence that contradicted this belief and showed that successive shots are independent events ([http://psych.cornell.edu/sites/default/files/Gilo.Vallone.Tversky.pdf](http://psych.cornell.edu/sites/default/files/Gilo.Vallone.Tversky.pdf)). This paper started a great controversy that continues to this day, as you can see by Googling *hot hand basketball*.

We do not expect to resolve this controversy today. However, in this activity we'll apply one approach to answering questions like this. The goals for the remainder of this workbook are to (1) think about the effects of independent and dependent events, (2) learn how to simulate shooting streaks in `R`, and (3) to compare a simulation to actual data in order to determine if the hot hand phenomenon appears to be real.

###

```{r hh-binomial-assumptions, echo = FALSE}
question_checkbox(
  "Which of the following assumptions must be made if we are to simulate a set of basketball shots as a binomial experiment?",
  answer("Every shot has two possible outcomes -- a 'hit' or a 'miss'", correct = TRUE),
  answer("If a player 'hits' a shot, then the likelihood that they make their next shot goes up"),
  answer("Shots are independent -- that is, the hot hand phenomenon does not exist", correct = TRUE),
  answer("We must simulate at least 50 shots"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

Our investigation will focus on the performance of one player: Kobe Bryant of the Los Angeles Lakers. His performance against the Orlando Magic in the 2009 NBA finals earned him the title *Most Valuable Player* and many spectators commented on how he appeared to show a hot hand. Let's load some data from those games and look at the first several rows.

```{r load-data, echo = FALSE, eval=TRUE}
head(kobe)
```

In this data frame, every row records a shot taken by Kobe Bryant. If he hit the shot (made a basket), a hit, `H`, is recorded in the column named `basket`, otherwise a miss, `M`, is recorded.

###

Just looking at the string of hits and misses, it can be difficult to gauge whether or not it seems like Kobe was shooting with a hot hand. One way we can approach this is by considering the belief that hot hand shooters tend to go on shooting streaks. For this lab, we define the length of a shooting streak to be the *number of consecutive baskets made until a miss occurs*.

For example, in Game 1 Kobe had the following sequence of hits and misses from his nine shot attempts in the first quarter:

\[ \textrm{H M | M | H H M | M | M | M} \]

To verify this run the command `kobe$basket[1:9]` using the code block below. Remember that this accesses the `basket` column of the `kobe` data frame, and the `[1:9]` says that we want entry 1 through entry 9 of that column.

```{r first9, echo = FALSE, exercise = TRUE}

```

###

Within the nine shot attempts, there are six streaks, which are separated by a "|" above. Their lengths are one, zero, two, zero, zero, zero (in order of occurrence).

```{r streak0, echo = FALSE}
quiz(
  question_radio(
    "What does a streak length of 1 mean, i.e. how many hits and misses are in a streak of 1?",
    answer("One hit"),
    answer("One hit and one miss", correct = TRUE),
    answer("One hit and any number of misses"),
    answer("Just a miss"),
    answer("Any number of misses"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "What does a streak of length 0 mean?",
    answer("One hit"),
    answer("One hit and one miss"),
    answer("One hit and any number of misses"),
    answer("Just a miss", correct = TRUE),
    answer("Any number of misses"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

In this workbook you have access to a custom function `calc_streak`, which was loaded in with the data. This function can be used to calculate the lengths of all shooting streaks and then look at the distribution.

Store the result of `calc_streak(kobe$basket)` to a new object called `kobe_streak`and then call `kobe_streak` if you want to see the streak lengths printed out.
```{r calc-streak-kobe, echo = FALSE, exercise = TRUE}

```

<div id = "calc-streak-kobe-hint">
Remember that assignment can be made with the `<-` operator.
</div>

###

Now use `ggplot` to make a barplot of `kobe_streak`.

```{r plot-streak-kobe, echo = FALSE, exercise = TRUE}

```

<div id = "plot-streak-kobe-hint">
Since `kobe_streak` is a vector rather than a data frame you can leave the *data* argument for `ggplot()` empty. That is, your code should start with `ggplot() +`
Be sure to check back on the Data Visualization workbook if you need a refresher on plotting.
</div>

###

Note that instead of making a histogram, we chose to make a bar plot from a table of the streak data. A bar plot is preferable here since our variable is discrete -- counts -- instead of continuous.

###

```{r describe-streaks, echo = FALSE}
quiz(
  question_radio(
    "Describe the distribution of Kobe's streaks from the 2009 NBA finals",
    answer("The distribution is uniform"),
    answer("The distribution is symmetric"),
    answer("The distribution is skewed left"),
    answer("The distribution is skewed right", correct = TRUE),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "What was Kobe's typical streak length?",
    answer("0", correct = TRUE),
    answer("1"),
    answer("2"),
    answer("3"),
    answer("4"),
    answer("5"),
    allow_retry = TRUE
  ),
  question_radio(
    "How long was Kobe's longest streak?",
    answer("0"),
    answer("1"),
    answer("2"),
    answer("3"),
    answer("4", correct = TRUE),
    answer("5"),
    allow_retry = TRUE
  )
)
```

## Compared to What?

We've shown that Kobe had some long shooting streaks, but are they long enough to support the belief that he had hot hands? What can we compare them to?

To answer these questions, let's return to the idea of *independence*. Two processes are independent if the outcome of one process doesn't effect the outcome of the second. If each shot that a player takes is an independent process, having made or missed your first shot will not affect the probability that you will make or miss your second shot.

###

A shooter with a hot hand will have shots that are *not* independent of one another. Specifically, if the shooter makes his first shot, the hot hand model says he will have a *higher* probability of making his second shot.

Let's suppose for a moment that the hot hand model is valid for Kobe. During his career, the percentage of time Kobe makes a basket (i.e. his shooting percentage) is about 45%, or in probability notation,

$$\mathbb{P}\left[\textrm{shot 1 = H}\right] = 0.45$$

If he makes the first shot and has a hot hand (*not* independent shots), then the probability that he makes his second shot would go up to, we might assume something like 60%,

$$\mathbb{P}\left[\textrm{shot 2 = H} \, | \, \textrm{shot 1 = H}\right] = 0.60$$
###

As a result of these increased probabilites, you'd expect Kobe to have longer streaks. Compare this to the skeptical perspective where Kobe does *not* have a hot hand, where each shot is independent of the next. If he hit his first shot, the probability that he makes the second is still 0.45.

$$\mathbb{P}\left[\textrm{shot 2 = H} \, | \, \textrm{shot 1 = H}\right] = 0.45$$

###

In other words, making the first shot did nothing to effect the probability that he'd make his second shot. If Kobe's shots are independent, then he'd have the same probability of hitting every shot regardless of his past shots: 45%.

###

Now that we've phrased the situation in terms of independent shots, let's return to the question: how do we tell if Kobe's shooting streaks are long enough to indicate that he has hot hands? We can compare his streak lengths to someone without hot hands: an independent shooter. 

## Simulations in R

We'll come back to Kobe shortly, but for now let's think about how we might simulate an independent shooter. While we don't have any data from a shooter we know to have independent shots, that sort of data is very easy to simulate in R. In a simulation, you set the ground rules of a random process and then the computer uses random numbers to generate an outcome that adheres to those rules. For example, the following code block is set up to simulate a single flip of a fair coin. Run it a few times and see what happens -- then edit the code to simulate 10 flips at a time, or a hundred, or maybe even a hundred-thousand! The computer can flip 100,000 coins much faster than you and I could ever hope to. 

```{r head-tail, exercise = TRUE, exercise.eval = FALSE}
sample(c("heads", "tails"), size = 1, replace = TRUE)
```

The first argument `c("heads", "tails")` represents the *outcomes* and can be thought of as a hat with two slips of paper in it: one slip says `heads` and the other says `tails`. The function `sample` draws one slip from the hat and tells us if it was a head or a tail. 

###

As you've discovered, if you wanted to simulate flipping a fair coin 100 times, you could either run the `sample()` function 100 times or, more simply, adjust the `size` argument, which governs how many samples to draw (the `replace = TRUE` argument indicates we put the slip of paper back in the hat before drawing again). Use the code block below to simulate 100 flips of a fair coin, storing the results in an object called `sim_fair_coin`. Then use `table(sim_fair_coin)` to compute a frequency table for the outcomes -- what do you notice?

```{r sim-fair-coin, exercise = TRUE}

```

###

Since there are only two elements in the *outcomes*, the probability that we "flip" a coin and it lands heads is 0.5. Say we're trying to simulate an unfair coin that we know only lands heads 20% of the time. We can adjust for this by adding an argument called `prob`, which provides a vector of two probability weights. Execute the following code to simulate this unfair coin.

```{r sim-unfair-coin, exercise = TRUE, exercise.eval = FALSE}
sim_unfair_coin <- sample(c("heads", "tails"), size = 100, replace = TRUE, prob = c(0.2, 0.8))
table(sim_unfair_coin)
```

###

`prob=c(0.2, 0.8)` indicates that for the two elements in the *outcomes* vector, we want to select the first one, `heads`, with probability 0.2 and the second one, `tails` with probability 0.8. Another way of thinking about this is to think of the outcome space as a bag of 10 chips, where 2 chips are labeled "head" and 8 chips "tail". Therefore at each draw, the probability of drawing a chip that says "head"" is 20%, and "tail" is 80%.

## Simulating the Independent Shooter

Simulating a basketball player who has independent shots uses the same mechanism that we use to simulate a coin flip. To simulate a single shot from an independent shooter with a shooting percentage of 50% we type,

```{r sim-basket, eval=FALSE, echo = TRUE}
sim_basket <- sample(c("H", "M"), size = 1, replace = TRUE)
```

###

To make a valid comparison between Kobe and our simulated independent shooter, we need to align both their shooting percentage and the number of attempted shots. Use the code block below to help you find the number of shots we need to simulate.

```{r find-num-shots, exercise = TRUE, echo = FALSE}

```

<div id = "find-num-shots-hint">
Use `dim(kobe)` to find the dimensions of the Kobe data frame. Remember that each row corresponds to a shot taken by Kobe.
</div>

```{r changing-sample-sim, echo = FALSE}
quiz(
  question_radio(
    "How many shots did Kobe Bryant take during the 2009 NBA finals?",
    answer("0"),
    answer("133", correct = TRUE),
    answer("76"),
    answer("6"),
    answer("45"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "What should the shooting percentage be for our simulated random independent shooter?",
    answer("50%"),
    answer("45%", correct = TRUE),
    answer("60%"),
    answer("40%"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

Use your answers from the previous questions, our discussion about simulated coin flipping, and the code block below to simulate an independent shooter which we can compare to Kobe Bryant. Store the resuts of your simulated shots in an object called `sim_basket`.

```{r sim-independent, echo = FALSE, exercise = TRUE}

```

```{r sim-independent-solution}
sim_basket <- sample(c("H", "M"), size = 133, prob = c(0.45, 0.55), replace = TRUE)
```

```{r sim-independent-check}
grade_code()
```

###

Note that we've named the new vector `sim_basket`, the same name that we gave to the previous vector reflecting a shooting percentage of 50%. In this situation, R overwrites the old object with the new one, so always make sure that you don't need the information in an old vector before reassigning its name.

###

With the results of the simulation saved as `sim_basket`, we have the data necessary to compare Kobe to our independent shooter. We can look at Kobe's data alongside our simulated data. Execute the following code cell to see Kobe's actual shots taken as well as our simulated shooter's shots. I've created a new version of the `sim_basket` object so that we will all see the same simulated data.

```{r compare-basket-setup}
set.seed(8)
sim_basket <- sample(c("H", "M"), size = 133, replace = TRUE)
sim_streak <- calc_streak(sim_basket)
```

```{r compare-basket, exercise = TRUE}
kobe$basket
sim_basket
```

###

Both data sets represent the results of 133 shot attempts, each with the same 
shooting percentage of 45%. We know that our simulated data is from a shooter 
that has independent shots. That is, we know the simulated shooter does not have
a hot hand.

###

Use `calc_streak` function in the code block below to compute the streak lengths of `sim_basket` and store the result in an object called `sim_streak`.

```{r sim-streak, exercise = TRUE, exercise.setup = "compare-basket-setup"}

```

```{r sim-streak-solution, exercise.setup = "compare-basket-setup"}
sim_streak <- calc_streak(sim_basket)
```

```{r sim-streak-check}
grade_code()
```

###

Be prepared to discuss the following questions in our small groups this week. You can use the code blocks below each question to help you answer.

1. Describe the distribution of streak lengths. What is the typical streak length for this simulated independent shooter with a 45% shooting percentage? How long is the simulated player's longest streak of baskets in 133 shots?

```{r discuss-q1, echo = FALSE, exercise = TRUE, exercise.setup = "compare-baseket-setup"}

```

2. If you were to run the simulation of the independent shooter a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning.

```{r discuss-q2, echo = FALSE, exercise = TRUE, exercise.setup = "compare-baseket-setup"}

```

3. How does Kobe Bryant's distribution of streak lengths compare to the distribution of streak lengths for the simulated shooter? Using this comparison, do you have evidence that the hot hand model fits Kobe's shooting patterns? Explain.

```{r discuss-q3, echo = FALSE, exercise = TRUE, exercise.setup = "compare-baseket-setup"}

```

<div id="license">
This is a derivative of a product of OpenIntro that is released under a 
[Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
The original lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel
from a lab written by Mark Hansen of UCLA Statistics.
</div>
