---
title: "Topic 17: Inference for Numerical Data"
tutorial:
  id: "InferenceNumericalLab.Topic17.AppliedStats"
  version: 1.0
output: 
  learnr::tutorial:
    progressive: TRUE
    
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(eval = FALSE)
#library(tidyverse)
library(dplyr)
library(ggplot2)
library(learnr)
library(gradethis)
library(skimr)

load("www/nc.RData")

tutorial_options(exercise.timelimit = 60, exercise.checker = gradethis::grade_learnr)

#knitr::opts_chunk$set(eval = FALSE)
```

## Inference for Numerical Data (Lab)

In this workbook, we consider multiple variables measured on a random sample of pregnancies in 2004. During that year, the state of North Carolina released a large data set containing information on births recorded in this state. This data set is useful to researchers studying the relation between habits and practices of expectant mothers and the birth of their children. We will work with a random sample of observations from this data set. The data has been loaded into this workbook as `nc`.

###

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab written by the faculty and TAs of UCLA Statistics.
</div>

###

In the `nc` dataset we have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | ----------------------------------------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

###

Use the code block below to answer the questions that follow.
```{r nc-cases-questions, exercise = TRUE}

```

```{r nc-cases, echo = FALSE}
quiz(
  question_radio(
    "What are the cases in this data set?",
    answer("Each case corresponds to a pregnancy.", correct = TRUE),
    answer("Each case corresponds to a woman."),
    answer("Each case corresponds to a doctor visit."),
    answer("Each case corresponds to a baby."),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "How many cases are there in our sample?",
    answer("963"),
    answer("13"),
    answer("246"),
    answer("1,000", correct = TRUE),
    answer("1,387"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

As a first step in the analysis, we should consider summaries of the data. This can be done using base-R's `summary()` command or using `skim()` from the `skimr` package. Call one of these functions on the `nc` data set below and use the output to help you answer the questions that follow.

```{r nc-summary, exercise = TRUE}

```

###

As you review the variable summaries, consider which variables are categorical and which are numerical. For numerical variables, are there outliers? If you aren't sure or want to take a closer look at the data, make a graph. You can use the code block below to experiment.
```{r nc-explore-num-cat-out, exercise = TRUE}

```

```{r nc-explore-num-cat-out-questions, echo = FALSE}
question_checkbox(
  "Which of the variables in the `nc` dataset are numerical? Check all that apply.",
  answer("`fage`", correct = TRUE),
  answer("`mage`", correct = TRUE),
  answer("`mature`"),
  answer("`weeks`", correct = TRUE),
  answer("`premie`"),
  answer("`visits`", correct = TRUE),
  answer("`marital`"),
  answer("`gained`", correct = TRUE),
  answer("`weight`", correct = TRUE),
  answer("`lowbirthweight`"),
  answer("`gender`"),
  answer("`habit`"),
  answer("`whitemom`"),
  allow_retry = TRUE
)
```

###

Consider the possible relationship between a mother's smoking habit and the weight of her baby. Plotting the data is a useful first step because it helps us quickly visualize trends, identify strong associations, and develop research questions.

Make a side-by-side boxplot of `habit` and `weight`. What does the plot highlight about the relationship between these two variables?
```{r plot-habit-weight, exercise = TRUE}

```

```{r plot-habit-weight-solution}
ggplot(data = nc) + geom_boxplot(mapping = aes(x = habit, y = weight))
```

```{r plot-habit-weight-check}
grade_code()
```

###

The box plots show how the medians of the two distributions compare, but we can also compare the means of the distributions. We can use `dplyr`'s `group_by()` to group our data according to the `habit` variable, and then `summarize()` the `weight` variable using the `mean()`. Consider the following code, guess what the result will look like, then run the code to see if you were right. What does the output tell you?
```{r by-means, exercise = TRUE}
nc %>%
  group_by(habit) %>%
  summarize(avg_birth_weight = mean(weight))
```

###

There is an observed difference in average birth weights, but is this difference statistically significant? In order to answer this question we will conduct a hypothesis test.

## Inference

Check if the conditions necessary for inference are satisfied. Note that you will need to obtain sample sizes to check the conditions. You can compute the group size by simply using the `count()` function.

```{r check-inference-conditions-habit, exercise = TRUE}

```

```{r check-inference-conditions-habit-solution}
nc %>%
  count(habit)
```

```{r check-inference-conditions-habit-check}
grade_code()
```

```{r inference-habit-conditions-questions, echo = FALSE}
quiz(
  question_radio(
    "How many groups are being considered?",
    answer("one"),
    answer("two", correct = TRUE),
    answer("more than two"),
    allow_retry = TRUE
  ),
  question_checkbox(
    "Which of the following are the groups?",
    answer("Pregnancies with mothers who are smokers", correct = TRUE),
    answer("Pregnancies with mothers who are non-smokers", correct = TRUE),
    answer("Pregnancies resulting in premature births"),
    answer("Pregnancies not resulting in premature births"),
    answer("Pregnancies resulting in babies of low birthweight"),
    answer("Pregnancies resulting in babies not having low birthweight"),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question_radio(
    "Are the conditions for inference satisfied?",
    answer("Yes, the group sizes are large enough to overcome any skew in the population distributions.", correct = TRUE),
    answer("No, the groups are too small."),
    answer("No, the group sizes are imbalanced."),
    answer("No, there is skew in the data."),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

###

Write the hypotheses for testing if the average weights of babies born to smoking and non-smoking mothers are different.
```{r hypotheses-sVns, echo = FALSE}
question(
  "The hypotheses for the test are:",
  answer("$\\begin{array}{ll} H_0: & \\mu_{\\text{smoker}} = \\mu_{\\text{non-smoker}}\\\\ H_a: & \\mu_{\\text{smoker}} \\neq \\mu_{\\text{non-smoker}}\\end{array}$", correct = TRUE),
  answer("$\\begin{array}{ll} H_0: & \\mu_{\\text{smoker}} = \\mu_{\\text{non-smoker}}\\\\ H_a: & \\mu_{\\text{smoker}} > \\mu_{\\text{non-smoker}}\\end{array}$"),
  answer("$\\begin{array}{ll} H_0: & \\mu_{\\text{smoker}} = \\mu_{\\text{non-smoker}}\\\\ H_a: & \\mu_{\\text{smoker}} < \\mu_{\\text{non-smoker}}\\end{array}$"),
  answer("$\\begin{array}{ll} H_0: & p_{\\text{smoker}} = p_{\\text{non-smoker}}\\\\ H_a: & p_{\\text{smoker}} = p_{\\text{non-smoker}}\\end{array}$"),
  answer("$\\begin{array}{ll} H_0: & p_{\\text{smoker}} = p_{\\text{non-smoker}}\\\\ H_a: & p_{\\text{smoker}} < p_{\\text{non-smoker}}\\end{array}$"),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

###

Next, we remind ourselves of the `inference()`, that we will use for conducting hypothesis tests and constructing confidence intervals. 

```{r inf-weight-habit-ht, eval=FALSE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

As a reminder, we'll go through the arguments of this custom function. The first argument is `y`, which is the response variable that we are interested in: `nc$weight`. The second argument is the explanatory variable, `x`, which is the variable that splits the data into two groups, smokers and non-smokers: `nc$habit`. The third argument, `est`, is the parameter we're interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a confidence interval (`"ci"`). When performing a hypothesis test, we also need to supply the `null` value, which in this case is `0`, since the null hypothesis sets the two population means equal to each other. The `alternative` hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` of inference can be `"theoretical"` or `"simulation"` based.

###

You should feel comfortable using your knowledge of inference to describe what this function must do for the parameter selections `type = "ht"` and for `type = "ci"`.

###

The initial `inference()` example code appears below. Think about the code and what you expect the output should look like. Run the code and check your expectations of the output. Once you understand the output, change the `type` argument to `"ci"` to construct and record a confidence interval for the difference between the weights of babies born to smoking and non-smoking mothers. Be sure to get rid of any additional unnecessary arguments as well.
```{r first-inference-run, exercise = TRUE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, alternative = "twosided", method = "theoretical")
```

```{r first-inference-run-solution}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", method = "theoretical")
```

```{r first-inference-run-check}
grade_code()
```

###

The results you see show that by default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$) because $\tt{R}$ orders levels by alphabetical order. We can easily change this order by using the `order` argument. Inspect the code below to identify what has been added and then run it to see the result.

```{r inference-ci-ordered, exercise = TRUE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", method = "theoretical", order = c("smoker","nonsmoker"))
```

## On your own

Since you've worked through this workbook, you should now feel prepared to answer the following. Use the code blocks to check your understanding.

1. Calculate a 95% confidence interval for the average length of pregnancies (`weeks`) and interpret it in context. Note that since you're doing inference on a single population parameter, there is no explanatory variable, so you can omit the `x` variable from the function.
```{r own-prob-1, exercise = TRUE}

```

2. Calculate a new confidence interval for the same parameter at the 90% confidence level. You can change the confidence level by adding a new argument to the function: `conflevel = 0.90`.
```{r own-prob-2, exercise = TRUE}

```

3. Conduct a hypothesis test evaluating whether the average weight gained by younger mothers is different than the average weight gained by mature mothers.
```{r own-prob-3, exercise = TRUE}

```

4. Now, a non-inference task: Determine the age cutoff for younger and mature mothers. Use a method of your choice, and explain how your method works.
```{r own-prob-4, exercise = TRUE}

```

5. Pick a pair of numerical and categorical variables and come up with a research question evaluating the relationship between these variables. Formulate the question in a way that it can be answered using a hypothesis test and/or a confidence interval. Answer your question using the `inference` function, report the statistical results, and also provide an explanation in plain language.
```{r own-prob-5, exercise = TRUE}

```

## Submit

```{r context="server"}
learnrhash::encoder_logic(strip_output = TRUE)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = shiny::div(
    "If your instructor is asking you to submit evidence of your completed notebook using a hash code, you may generate that code below. Use the clipboard icon to ensure that you are copying the entire hash."
    )
  )
```

<br/>

<div id="license">
This lab is a product of OpenIntro that is released under a [Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab written by the faculty and TAs of UCLA Statistics.
</div>
